The Neuro Bureau ADHD-200 Preprocessed Repository

### 2016-05-26 Bellec_ADHD200_2016.tex
Added links to the "consolidated phenotypics".

With the exception of a few datasets that were missing data, and one dataset that was corrupted, all of the ADHD-200 Sample was processed and released by Athena, regardless of data quality. This was done to accomodate differing opinions as to what qualifies as usable data, and to provide poor quality data that may be used by others to develop methods that are robust to noise. Files containing the six motion parameters for each rs-fMRI data and anatomical and mean EPI images in template space have been included in the release to enable users to determine high-motion data or poor registrations. 
\par
An online report\footnote{\url{http://simexp.github.io/adhd200_qc_athena/index.html}} of registration quality was generated for all subjects. The report includes concurrent visualization of the individual structural images, after non-linear registration, together with the target brain template (MNI pediatric), as well as a concurrent visualization of the individual structural and BOLD images together, again after non-linear registration. All these images were visually reviewed, and registration that presented severe misregistration or image quality problems were marked as ``fail'' for QC ($7\%$ failure rate). The outcome of the visual quality assessment for registered images has been shared on NITRC in a consolidated tabular-separated (.tsv) spreadsheet that also includes all phenotypic variables. The .tsv QC spreadsheet also includes a summary measure of motion amplitude for each dataset, that can be used to exclude subjects above a certain threshold. 
\par
Additional quality metrics derived from the preprocessed s-MRI, GM masks, mean rs-fMRI images, fALFF maps, and FC maps were also included to help with the QC process. For each data type, registered in stereotaxic space,  a mean and standard deviation image was calculated from all of the scans from all of the subjects. These images were used to perform a voxelwise z-score transformation on each data type for each subject $z_i = \frac{(v_i-m_i)}{\sigma_i}$. The absolute values of these z-scores were thresholded at 3, and summed across voxels for a quality score for that image. The higher the resulting sum, the larger the number of voxels a image has that are $\geq 3$ and the more likely the images are outliers. Using this metric it is possible to rank images, and allows for a more directed search for poor quality scans. These metrics are distributed in participant-specific text files along with the BASH scripts that implement the procedure.

@@

Outputs of the NIAK pipeline were subjected to a careful visual inspection and the QC reports, along with head motion statistics, are available on the NIAK description page \footnote{\url{http://www.nitrc.org/plugins/mwiki/index.php/neurobureau:NIAKPipeline#Quality_control_of_the_preprocessing_-_Training_dataset}}. Estimates of the maximum motion (translation and rotation) between consecutive functional volumes for each rs-fMRI dataset were inspected to categorize the datasets as containing minimal ($<$1mm or degree), moderate (2 to 3 mm or degrees) or severe motion ($>$3 mm or degrees). The individual brain registration of the NIAK pipeline were visually inspected using online QC reports\footnote{\url{simexp.github.io/adhd200_qc_niak/}}, similar to those generated for the Athena pipeline. When substandard registration outcomes were identified, a parameter controlling the non-uniformity correction of the s-MRI was adjusted and the analysis was repeated until the coregistration results were satisfactory. Satistfactory results could not be achived with some datasets and have been indicated as ``Fail'' for QC ($5.2\%$ failure rate) in the .tsv spreadsheet including QC assessments, available on NITRC. 
%%%
%%% - tables
%%%

@@

The ADHD-200 preprocessed has helped to expand the boundaries of the traditional neuroimaging community, with several publications in core engineering and statistics journals that do not routinely feature neuroimaging applications, such as Statistica Sinica \citep{Yang2015} or the Journal of the American Statistical Association \citep{Chen2015}, amongst others \citep[e.g.][]{Liang2012, Rangarajan2015, Rangarajan2014, Mahanand2013, Ji2011, Li2015, Li2013, Liu2012, Subramanian2013, Hou2015, Deshpande2015, She2014, He2013, Kong2013, Ahn2015}. In particular, the winning team of the ADHD-200 Global Competition was based at the Johns Hopkins Biostatistics Department and used ADHD-200 Preprocessed to develop their diagnostic algorithm \cite{Eloyan2012}. We also found interesting that a handful of methodological works used the ADHD-200 Preprocessed sample as one application in a series of benchmarks that could be as varied as positron emission tomography in Alzheimer's disease \citep{Yang2015}, the Yeast gene regulatory network \citep{She2014} or gene expression in brain tissues of patients with HIV-1 associated neurocognitive disorders \citep{Liu2012}. This suggests that fully processed, easy-to-access samples could dramatically accelerate the cross-polinization of disciplines.

>>>>

With the exception of a few datasets that were missing data, and one dataset that was corrupted, all of the ADHD-200 Sample was processed and released by Athena, regardless of data quality. This was done to accomodate differing opinions as to what qualifies as usable data, and to provide poor quality data that may be used by others to develop methods that are robust to noise. Files containing the six motion parameters for each rs-fMRI data and anatomical and mean EPI images in template space have been included in the release to enable users to determine high-motion data or poor registrations. 
\par
An online report\footnote{\url{http://simexp.github.io/adhd200_qc_athena/index.html}} of registration quality was generated for all subjects. The report includes concurrent visualization of the individual structural images, after non-linear registration, together with the target brain template (MNI pediatric), as well as a concurrent visualization of the individual structural and BOLD images together, again after non-linear registration. All these images were visually reviewed, and registration that presented severe misregistration or image quality problems were marked as ``fail'' for QC ($7\%$ failure rate). The outcome of the visual quality assessment for registered images has been shared on NITRC in a consolidated tabular-separated (.tsv) spreadsheet\footnote{\url{https://www.nitrc.org/frs/download.php/9024/adhd200_preprocessed_phenotypics.tsv}} that also includes all phenotypic variables. The .tsv QC spreadsheet also includes a summary measure of motion amplitude for each dataset, that can be used to exclude subjects above a certain threshold. 
\par
Additional quality metrics derived from the preprocessed s-MRI, GM masks, mean rs-fMRI images, fALFF maps, and FC maps were also included to help with the QC process. For each data type, registered in stereotaxic space,  a mean and standard deviation image was calculated from all of the scans from all of the subjects. These images were used to perform a voxelwise z-score transformation on each data type for each subject $z_i = \frac{(v_i-m_i)}{\sigma_i}$. The absolute values of these z-scores were thresholded at 3, and summed across voxels for a quality score for that image. The higher the resulting sum, the larger the number of voxels a image has that are $\geq 3$ and the more likely the images are outliers. Using this metric it is possible to rank images, and allows for a more directed search for poor quality scans. These metrics are distributed in participant-specific text files along with the BASH scripts that implement the procedure.

@@

Outputs of the NIAK pipeline were subjected to a careful visual inspection and the QC reports, along with head motion statistics, are available on the NIAK description page \footnote{\url{http://www.nitrc.org/plugins/mwiki/index.php/neurobureau:NIAKPipeline#Quality_control_of_the_preprocessing_-_Training_dataset}}. Estimates of the maximum motion (translation and rotation) between consecutive functional volumes for each rs-fMRI dataset were inspected to categorize the datasets as containing minimal ($<$1mm or degree), moderate (2 to 3 mm or degrees) or severe motion ($>$3 mm or degrees). The individual brain registration of the NIAK pipeline were visually inspected using online QC reports\footnote{\url{simexp.github.io/adhd200_qc_niak/}}, similar to those generated for the Athena pipeline. When substandard registration outcomes were identified, a parameter controlling the non-uniformity correction of the s-MRI was adjusted and the analysis was repeated until the coregistration results were satisfactory. Satisfactory results could not be achived with some datasets and have been indicated as ``Fail'' for QC ($5.2\%$ failure rate) in the .tsv spreadsheet including QC assessments, available on NITRC\footnote{\url{https://www.nitrc.org/frs/download.php/9024/adhd200_preprocessed_phenotypics.tsv}}. 
%%%
%%% - tables
%%%

@@

The ADHD-200 preprocessed has helped to expand the boundaries of the traditional neuroimaging community, with several publications in core engineering and statistics journals that do not routinely feature neuroimaging applications, such as Statistica Sinica \citep{Yang2015} or the Journal of the American Statistical Association \citep{Chen2015}, amongst others \citep[e.g.][]{Liang2012, Rangarajan2015, Rangarajan2014, Mahanand2013, Ji2011, Li2015, Li2013, Liu2012, Subramanian2013, Hou2015, Deshpande2015, She2014, He2013, Kong2013, Ahn2015}. In particular, the winning team of the ADHD-200 Global Competition was based at the Johns Hopkins Biostatistics Department and used ADHD-200 Preprocessed to develop their diagnostic algorithm \cite{Eloyan2012}. We also found interesting that a handful of methodological works used the ADHD-200 Preprocessed sample as one application in a series of benchmarks that could be as varied as positron emission tomography in Alzheimer's disease \citep{Yang2015}, the Yeast gene regulatory network \citep{She2014} or gene expression in brain tissues of patients with HIV-1 associated neurocognitive disorders \citep{Liu2012}. This suggests that fully processed, easy-to-access imaging samples could help validate general-purpose methods on a wider scope of applications.
